Modeling tabular icebergs coupled to an ocean model

### 2017-04-03 submission1/JAMES_iceberg_only.tex
Final commit after submission of JAMES paper.
I have added all movies to the suplimentary material. The submitted version of the manuscript and all figures are in the submission_1 subdirectoy.

\begin{abstract}
Large tabular icebergs calved from Antarctic ice shelves have long lifetimes (due to their large size), during which they drift across large distances, altering ambient ocean circulation, bottom-water formation, sea-ice formation and biological primary productivity in the icebergs' vicinity.
However, despite their importance, the current generation of ocean circulation models do not represent large tabular icebergs.
In this study we develop a novel framework to model large tabular icebergs submerged in the ocean. In this framework, tabular icebergs are represented by Lagrangian elements that drift in the ocean, and are held together and interact with each other via bonds.
%We describe the Lagrangian bonded iceberg model and discuss a number of the technical elements involved in its development. 
%We demonstrate the capabilities of the model by modeling a tabular iceberg drifting away an idealized ice shelf.} 
A break of these bonds allows the model to emulate calving events (i.e. detachment of a tabular iceberg from an ice shelf) and tabular icebergs breaking up into smaller pieces. Idealized simulations of a calving tabular iceberg, its drift, and its breakup, demonstrate capabilities of the developed framework.
\end{abstract}

@@

%As mentioned above, we simplify the KID model by choosing to not explicitly calculate the angular momentum of the ice elements, which is computationally expensive. Instead, we allow rotational motion of larger structures emerge as a consequence of bond orientation and collective motion \citep{Jakobsen2001}.
%

@@

%In this latter case, the KID can be categorized as a discrete element model (DEM). 
% in the sense that large structures are constructed from smaller independent Lagrangian elements.

@@

In addition to the external forces, the ice elements experience interactive forces due to the presence of other elements. Two types of interactive forces are included between elements. The first force is a repulsive force which is applied to elements to prevent them from overlapping the boundaries of the neighboring elements. The second interactive force is a force due to numerical `bonds', and is only applied if two elements are labelled as `bonded'. When two elements are bonded, each element feels an attractive force that prevents the elements from moving too far apart from one another.  The details of the interactive forces are provided in below. 
%The interactive forces between two bonded elements are defined such that in the absence of other forces the elements come to rest adjacent to one another, with no overlap of the iceberg areas. \textcolor{red}{???The last sentence is unclear???}

@@

The coupling from the ocean model to the iceberg model is straight forward: at every time step: the ocean mixed layer temperature, salinity, velocity and sea-ice concentration are passed from the ocean model to the iceberg model, to be used in the momentum and thermodynamic equations of the ice elements. 
Since tabular icebergs are explicitly resolved in the ocean, it is sufficient for each element to interact with ocean mixed layer only (i.e.: there is no need to manually embed icebergs into the ocean by integrating ocean fields over the icebergs' thickness, as suggested in \citet{Merino2016}).
Within the KID model, the ocean model fields are interpolated onto the Lagrangian grid using a bilinear interpolation scheme.% The KID is not sensitive to the chosen interpolation scheme.

%More care needs to be taken when passing fields back from the KID to the ocean model. 
%The coupling from the KID to ocean model is more complex.
The iceberg model influences the ocean by: (i) applying a pressure to the ocean surface, (ii) affecting the upper ocean by applying a no-slip boundary condition and frictional velocity beneath the ice, and (iii) imposing heat, salt and mass fluxes on the ocean, associated with ice melting.
Six fields are passed from the iceberg model to the ocean model: ice mass, ice area, frictional velocity, and heat, salt and mass fluxes. 
Fields in the iceberg model are aggregated from the Lagrangian elements to the Eulerian ocean grid before they are passed to the ocean model.

@@

The numerical calculation of the intersection between hexagons and the ocean grid is simplified by dividing the hexagon into 6 equilateral triangles. This method allows for the intersection to be found even when the hexagon is not aligned with the grid. %The KID can be run using hexagonal or rectangular elements.

@@

\subsubsection{Parallelization and halos}
The KID model runs on multiple processors in parallel (using the same grid decomposition as the ocean model). When elements move from an ocean cell on one processor to an ocean cell on a second processor, the memory has to be passed from one processor the next, added and removed to the appropriate lists and the memory has to be allocated and deallocated correctly. Element interactions across the edge of processors are handled using computational halos. A computational halo is a copy of the edge of a one processor which is appended to the edge of a second processor, so that the first processor can interact with the second processor during a time step. Before each time step, elements at the edges of each processor are copied onto the halos of adjacent processors so that they can be used in calculating the interactive forces. After each time step, these halos are emptied, and the process is repeated. These halo updates are one of the most computationally expensive parts of the iceberg model. Details of how the bonds are broken and reconnected across processor boundaries are provided in Appendix~C.

@@

%We demonstrate this capability in an experimental setup using an idealized ice shelf. For these experiments the KID is coupled to the MOM6 ocean model.
% This section contains a description of the experimental setup and numerical experiments performed. The results of these simulations are presented in Section 4.

 %In the examples below, the KID is firstly used in uncoupled mode to model the movement of freely drifting icebergs. Secondly, the coupled KID with static ice elements is used to model an idealized ice shelf cavity. Finally, the fully-coupled KID is used to model a tabular iceberg which calves off of this ice shelf.

%\subsection{Uncoupled simulation}
%We begin by demonstrating the KID uncoupled to the ocean. For this experiment, we use a rectangular domain, $L_{x}=1000$ km long and $L_{y}=500$ km wide. Land points are arranged in a semi-circle, in the west of the domain, as shown by the black dots in Figure \ref{fig:Uncoupled_sim}.  Three hundred ice elements are initialized with diameters uniformly sampled from the interval [0 40km], and positioned randomly at positions with $x \leq 300$km. These ice elements can be thought to represent interacting sea ice flows or icebergs. Three `tabular icebergs' are initialized by arranging ice elements in a staggered lattices with adjacent elements bonded together, as described in Section 2.4. The three tabular icebergs contain 25, 16 and 4 elements, respectively. Each of the elements that make up the tabular icebergs has a diameter of 40km. All elements have an initial thickness of 10m.
%
%The simulation is forced by an imposed westward ocean current of u=0.1m/s. Forces due to sea surface slope, atmospheric drag, Coriolis and sea ice drag are set to zero. The initial setup is shown in Figure \ref{fig:Uncoupled_sim}a.

@@

 \section{Experiment Setup}
The introduction of Lagrangian elements, numerical bonds and interpolation schemes between the Eularian and Lagrangian grids (discussed in Section 2) means that we now have the tools to model large tabular icebergs submerged in the ocean. We demonstrate this capability by simulating a tabular iceberg drifting away from an ice shelf in idealized setting. 
%This simulation is performed using the KID coupled to the MOM6 ocean model \citep{Hallberg2013}. 
%This section describes the experimental setup of these numerical experiments. The results are shown in Section 4.

@@

\subsection{Ocean model setup}
The iceberg model is coupled to the MOM6 ocean model \citep{Hallberg2013}. The ocean model configuration uses a vertical coordinate system which is a hybrid between a sigma-level and a z-level coordinate. In particular, model layers deform underneath the ice shelf as they would in a sigma-coordinate model, but collapse to zero thickness when they intersect with bottom topography, as they would in a z-level model. The coordinate system was achieved using ALE regridding-remapping scheme \citep{White2009}. The model uses a horizontal resolution of 2~km, and 72 vertical layers.
All simulations were repeated using the ocean model configured in isopycnal mode (results were similar and are not presented here).
%(select results are shown in the supplementary material).

@@

\subsection{Breaking bonds}
The numerical bonds in the KID model enable the tabular iceberg to retain its shape. This is demonstrated by comparing the Control simulation to an identical simulation where all numerical bonds have been removed (Figure \ref{fig:Drift_berg}, movie S5). In the bond-free simulation, the ice elements disperse and the calved iceberg quickly loses its original structure. This bond-free simulation does not adequately represent tabular iceberg, which can move long distances through the ocean as a cohesive unit. This result motivates the inclusion of bonds in the iceberg model, even though they are more computationally expensive than traditional point-iceberg models. 
%The model behavior in the bond-free simulations is similar to an ice-shelf disintegration event.%, suggesting that 
%Breaking multiple bonds simultaneously could be used to simulate ice shelf disintegration events. \textcolor{red}{??? either add something more or remove this sentence ???}

@@

The  wave radiation force $(\vec{F}_{R})$ is given by
\begin{equation}
\vec{F}_{R} = \frac{1}{2}\rho_{o} c_{r} g a \textrm{ min}(a,F)  \textrm{ } 2\frac{WL}{W+L} \frac{\vec{v}_{a}}{|\vec{v}_{a} |}
\end{equation}
where g is the acceleration due to gravity, a is the wave amplitude empirically related to the wind speed by  $a=0.010125|\vec{v}_{a} -  \vec{v}_{o}| $, and $c_{wd}$ is the wave drag coefficient defined as 
\begin{equation}
c_{wd}=0.06 \textrm{ min} \Bigg( \textrm{ max} \Bigg[0,\frac{L-L_{c}}{L_{t}-L_{c} } \Bigg]  , 1    \Bigg), 
\end{equation}
where $L_{w}=0.32|\vec{v}_{a} -  \vec{v}_{o}|^{2}$ is an empirical wave length, $L_{c}=0.125 L_{w}$ is the cutoff length, and  $L_{t}=0.25 L_{w}$ is the upper limit.

@@

\subsection{Melt rate parametrization}
As discussed in Section 2.5, unbounded ice elements in the KID decay according to parameterizations for iceberg decay typically used in iceberg drift models \citep{Martin2010}, while ice elements within larger ice structures have only a basal melt given by the three equation model \citep{Holland1999} .

For unbonded ice elements, the element thickness decays due to basal melt at a rate $M_{b}$, while the length and width of the elements decay as a result of melt due to wave erosion, $M_{e}$, and melt due to buoyant convection, $M_{v}$. Following \citet{Gladstone2001} and \citet{Martin2010}, the basal melt rate, wave erosion melt rate, and buoyant convection melt rate are parameterized by 
\begin{equation}
M_{b} = 0.58 | \vec{v} - \vec{v}_{0}|^{0.8} \frac{\tilde{T}_{0} - \tilde{T}}{L^{0.2}}
\end{equation}
\begin{equation}
M_{e} = \frac{1}{12} S_{s} \Bigg(  1 + \cos{[\pi A_{i}^{3}]}  \Bigg) \Bigg( \tilde{T}_{0} +2 \Bigg ),
\end{equation}
\begin{equation}
M_{v} =  \Bigg(7.62 \times 10^{-3}  \Bigg )   \tilde{T}_{0} + \Bigg( 1.29 \times 10^{-3} \Bigg)\tilde{T}_{0}^{2}.
\end{equation}
$\tilde{T}$ is the effective iceberg temperature and is set to $\tilde{T}=4^{\circ}$C, $\tilde{T}_{0}$ is the temperature at the top of the ocean,  $A_{i}$ is the sea-ice area fraction, and $S_{s}$ is the sea state, which is given by the Beaufort scale
\begin{equation}
S_{s} =  \frac{2}{3}  | \vec{u}_{a} - \vec{u}_{o} |^{\frac{1}{2}}  +\frac{1}{10}   | \vec{u}_{a} -\vec{u}_{o} |
\end{equation}
All three melt rates are in units of meters per day.

@@

%The point-particle iceberg melt rates and the 3 equation model melt rates are both calculated using the ocean mixed-layer properties, which are passed from the ocean model to the KID.

@@

\section{Appendix B}
\subsection{Modified Verlet Algorithm}
The KID uses a version velocity Verlet time-stepping algorithm, which has been modified to allow part of the forcing to be calculated implicitly. The traditional velocity Verlet algorithm is commonly used in molecular dynamics, as it is simple to implement, second order accurate and computationally efficient \citep{Swope1982, Omelyan2002}.
Here we modify the traditional scheme to allow for the drag forces to be modeled implicitly, which prevents large accelerations for element's whose mass approaches zero. 
To do this, we include both an implicit and explicit acceleration, $a=a^{exp} +a^{imp}$. The explicit acceleration, $a^{exp}$, includes all forcing terms which depend only on the previous time step and the current position, while the implicit acceleration, $a^{imp}$, includes forcing terms which depend on the velocity at the current time step (in particular the drag and Coriolis forces).

@@

Using a time step of $\Delta t$, and subscripts to denote the time step (so that $t_{n+1}= t_{n} + \Delta t$), the modified velocity Verlet scheme can be written as:\\
1) Calculate updated position: $ x_{n+1} = x_{n} + u_{n} \Delta t  + \frac{1}{2} \Delta t^{2} \Bigg(a^{exp}_{n} + a^{imp}_{n} \Bigg) $.\\
2) Calculate $a^{exp}_{n+1}$\\
% (which is a function of $x_{n+1}$, and other quantities measured at time $t_{n}$). \\
3) Calculate $a^{imp}_{n+1}$ and $u_{n+1} = u_{n} + \frac{\Delta t}{2} \Bigg(a^{exp}_{n} + a^{exp}_{n+1} \Bigg) +  (\Delta t) a^{imp}_{n+1}$\\
%  (which is a function of $u_{n+1}$, $x_{n+1}$, and other quantities measured at time $t_{n}$)
This scheme reduces to the traditional velocity Verlet when $a^{imp}$ is set to zero. Note that $a^{exp}_{n+1} = a^{exp}_{n+1}(x_{n+1}, t_{n})$ is an explicit function of  $x_{n+1}$ and other quantities evaluated at time $t_{n}$, while $a^{imp}_{n+1}=a^{imp}_{n+1}(u_{n+1}, x_{n+1}, t_{n})$ additionally depends on $u_{n+1}$, and needs to be solved implicitly. For this reason in step three, $a^{imp}_{n+1}$ and $u_{n+1}$ need to be solved simultaneously, as described in the next subsection.

@@

\section{Appendix C}
\textbf{Connecting bonds across processor boundaries}\\
Since the KID is parallelized across multiple processors, it often happens that two elements on different processes are bonded together. Keeping track of numerical bonds across processor boundaries requires a lot of book keeping. In this section we describe the how KID handles bonds across processor boundaries.

@@

\clearpage
\begin{table}[h!]
\begin{center}
\begin{tabular}{ |c c c c| }
\hline
Parameter & Symbol & Value &  Unit \\
\hline
Domain Length &  $L_{x}$ & 80  & km \\
Domain Width &  $L_{y}$ & 480  & km \\
Horizontal Resolution &  $\Delta x$ & 2  & km \\
Number of vertical layers &  $N_{l}$ & 72  & non-dim\\
Horizontal Viscosity &  $\nu_{H}$ & 6.0 & $m^{2}$ $s^{-1}$ \\
Diapycnal Viscosity &  $\nu_{V} $ & $10^{-3}$ & $m^{2}$ $s^{-1}$\\
Horizontal Diffusivity & $ \epsilon_{H}$ & 1.0 & $m^{2}$ $s^{-1}$ \\
Diapycnal Diffusivity &  $\epsilon_{V} $ & $5 \times 10^{-5}$ & $m^{2}$ $s^{-1}$ \\
Initial Surface Temperature &  $ T_{t}$& -1.9   & $ ^{o} C$\\
Initial Bottom Temperature &  $ T_{b}$&  1.0  & $ ^{o} C$\\
Initial Surface Salinity &  $ S_{t}$& 33.8   & psu\\
Initial Bottom Salinity &  $ S_{b}$&  34.7  & psu\\
Maximum Ocean depth &  $H_{ocean} $ & 720 & m \\
Relaxation Time of Sponge Layer &  $T_{sponge}$ & 0.1  & days\\
Length of Sponge Layer &  $L_{sponge}$ & 10  & km\\
%Background Frictional Velocity Below Ice &  $u^{*}_{bg}$ & 0.0  & 104 \\
%Drag Coefficient Below Ice&  c_{d} & $0.0025$  & (non dim)\\
%Class 9 &  $3.9\times 10^{11}$ & 1659 & 104 \\
%Time Step for Static Shelf Experiment & $dt_{Static}$ &  1000  & s\\
KID Time Step & $dt$ &  10  & s\\
Elastic interactive force spring constant & $ \kappa_{e} $ &  $10^{-5}$  & kg $\textrm{s}^{-2}$ \\
%Class 1 &  $8.8\times 10^{7}$ & 60  & 0.0026 & 40 & 2000 \\
%Class 2 &  $4.1\times 10^{8}$ & 100 & 0.0072 & 67 & 200\\
%Class 3 &  $3.3\times 10^{9}$ & 200  & 0.029 &133 & 50 \\
%Class 4 &  $1.8\times 10^{10}$& 350 &  0.12 &175  & 20 \\
%Class 5 &  $3.8\times 10^{10}$ & 500  & 0.18 &250 & 10\\
%Class 6 &  $7.5\times 10^{10}$ & 700   & 0.35 &250 & 5\\
%Class 7 &  $1.2\times 10^{11}$ & 900  & 0.56 &250  & 2 \\
%Class 8 &  $2.2\times 10^{11}$ & 1200  & 1.0&250 & 1 \\
%Class 9 &  $3.9\times 10^{11}$ & 1600 & 1.8&250 & 1  \\
%Class 10 & $7.4\times 10^{11}$ &  2200 & 3.5 &250  &1\\
\hline
\end{tabular}
\caption{Parameters used in the model. The ocean model parameters are as described in \citet{Asay-Davis2016}}.
% Note that the areas of the iceberg classes (calculated from the iceberg lengths and masses given in \citep{Gladstone2001}) are not monotonically increasing.
\label{table:parameters2}
\end{center}
\end{table}

@@

%\begin{figure}
%\begin{center}
%\includegraphics[width=0.99\textwidth]{/Users/alon/Desktop/files/Icebergs_clusters/Towards_Publication/Tech_paper/Github_stuff/Tech-paper/Figures/snapshots_ALE_z_Mixed_Melt_Collapse_spread_area.png}
%\caption{ {Snapshots of the fraction of ice cover in the KID tabular iceberg calving simulation. Snapshots are taken (a) 7, (b) 15, and (c) 30 days after calving. The dashed line in panel (c) shows the location of the vertical transects shown in Figures \ref{fig:Velocity_section_Collapse} and \ref{fig:Temperature_section_Collapse}}. \label{fig:Spread_area}}
%\end{center}
%\end{figure}
 %\clearpage

>>>>

\begin{abstract}
Large tabular icebergs calved from Antarctic ice shelves have long lifetimes (due to their large size), during which they drift across large distances, altering ambient ocean circulation, bottom-water formation, sea-ice formation and biological primary productivity in the icebergs' vicinity.
However, despite their importance, the current generation of ocean circulation models do not represent large tabular icebergs.
In this study we develop a novel framework to model large tabular icebergs submerged in the ocean. In this framework, tabular icebergs are represented by Lagrangian elements that drift in the ocean, and are held together and interact with each other via bonds.
A break of these bonds allows the model to emulate calving events (i.e. detachment of a tabular iceberg from an ice shelf) and tabular icebergs breaking up into smaller pieces. Idealized simulations of a calving tabular iceberg, its drift, and its breakup, demonstrate capabilities of the developed framework.
\end{abstract}

@@

%As mentioned above, we simplify the iceberg model model by choosing to not explicitly calculate the angular momentum of the ice elements, which is computationally expensive. Instead, we allow rotational motion of larger structures emerge as a consequence of bond orientation and collective motion \citep{Jakobsen2001}.
%

@@

%In this latter case, the model can be categorized as a discrete element model (DEM). 
% in the sense that large structures are constructed from smaller independent Lagrangian elements.

@@

In addition to the external forces, the ice elements experience interactive forces due to the presence of other elements. Two types of interactive forces are included between elements. The first force is a repulsive force which is applied to elements to prevent them from overlapping the boundaries of the neighboring elements. The second interactive force is a force due to numerical `bonds', and is only applied if two elements are labelled as `bonded'. When two elements are bonded, each element feels an attractive force that prevents the elements from moving too far apart from one another.  The details of the interactive forces are provided in below.

@@

The coupling from the ocean model to the iceberg model is straight forward: at every time step: the ocean mixed layer temperature, salinity, velocity and sea-ice concentration are passed from the ocean model to the iceberg model, to be used in the momentum and thermodynamic equations of the ice elements. 
Since tabular icebergs are explicitly resolved in the ocean, it is sufficient for each element to interact with ocean mixed layer only (i.e.: there is no need to manually embed icebergs into the ocean by integrating ocean fields over the icebergs' thickness, as suggested in \citet{Merino2016}).
Within the iceberg model, the ocean model fields are interpolated onto the Lagrangian grid using a bilinear interpolation scheme.% The model is not sensitive to the chosen interpolation scheme.

The iceberg model influences the ocean by: (i) applying a pressure to the ocean surface, (ii) affecting the upper ocean by applying a no-slip boundary condition and frictional velocity beneath the ice, and (iii) imposing heat, salt and mass fluxes on the ocean, associated with ice melting.
Six fields are passed from the iceberg model to the ocean model: ice mass, ice area, frictional velocity, and heat, salt and mass fluxes. 
Fields in the iceberg model are aggregated from the Lagrangian elements to the Eulerian ocean grid before they are passed to the ocean model.

@@

The numerical calculation of the intersection between hexagons and the ocean grid is simplified by dividing the hexagon into 6 equilateral triangles. This method allows for the intersection to be found even when the hexagon is not aligned with the grid.

@@

\subsubsection{Parallelization and halos}
The iceberg model runs on multiple processors in parallel (using the same grid decomposition as the ocean model). When elements move from an ocean cell on one processor to an ocean cell on a second processor, the memory has to be passed from one processor the next, added and removed to the appropriate lists and the memory has to be allocated and deallocated correctly. Element interactions across the edge of processors are handled using computational halos. A computational halo is a copy of the edge of a one processor which is appended to the edge of a second processor, so that the first processor can interact with the second processor during a time step. Before each time step, elements at the edges of each processor are copied onto the halos of adjacent processors so that they can be used in calculating the interactive forces. After each time step, these halos are emptied, and the process is repeated. These halo updates are one of the most computationally expensive parts of the iceberg model. Details of how the bonds are broken and reconnected across processor boundaries are provided in Appendix~C.

@@

%We demonstrate this capability in an experimental setup using an idealized ice shelf. For these experiments the iceberg model is coupled to the MOM6 ocean model.
% This section contains a description of the experimental setup and numerical experiments performed. The results of these simulations are presented in Section 4.

 %In the examples below, the iceberg mode is firstly used in uncoupled mode to model the movement of freely drifting icebergs. Secondly, the coupled iceberg model with static ice elements is used to model an idealized ice shelf cavity. Finally, the fully-coupled iceberg mode is used to model a tabular iceberg which calves off of this ice shelf.

%\subsection{Uncoupled simulation}
%We begin by demonstrating the iceberg mode uncoupled to the ocean. For this experiment, we use a rectangular domain, $L_{x}=1000$ km long and $L_{y}=500$ km wide. Land points are arranged in a semi-circle, in the west of the domain, as shown by the black dots in Figure \ref{fig:Uncoupled_sim}.  Three hundred ice elements are initialized with diameters uniformly sampled from the interval [0 40km], and positioned randomly at positions with $x \leq 300$km. These ice elements can be thought to represent interacting sea ice flows or icebergs. Three `tabular icebergs' are initialized by arranging ice elements in a staggered lattices with adjacent elements bonded together, as described in Section 2.4. The three tabular icebergs contain 25, 16 and 4 elements, respectively. Each of the elements that make up the tabular icebergs has a diameter of 40km. All elements have an initial thickness of 10m.
%
%The simulation is forced by an imposed westward ocean current of u=0.1m/s. Forces due to sea surface slope, atmospheric drag, Coriolis and sea ice drag are set to zero. The initial setup is shown in Figure \ref{fig:Uncoupled_sim}a.

@@

 \section{Experiment Setup}
The introduction of Lagrangian elements, numerical bonds and interpolation schemes between the Eularian and Lagrangian grids (discussed in Section 2) means that we now have the tools to model large tabular icebergs submerged in the ocean. We demonstrate this capability by simulating a tabular iceberg drifting away from an ice shelf in idealized setting. 
%This simulation is performed using the iceberg model coupled to the MOM6 ocean model \citep{Hallberg2013}. 
%This section describes the experimental setup of these numerical experiments. The results are shown in Section 4.

@@

\subsection{Ocean model setup}
The KID model is coupled to the MOM6 ocean model \citep{Hallberg2013}. The ocean model configuration uses a vertical coordinate system which is a hybrid between a sigma-level and a z-level coordinate. In particular, model layers deform underneath the ice shelf as they would in a sigma-coordinate model, but collapse to zero thickness when they intersect with bottom topography, as they would in a z-level model. The coordinate system was achieved using ALE regridding-remapping scheme \citep{White2009}. The model uses a horizontal resolution of 2~km, and 72 vertical layers.
All simulations were repeated using the ocean model configured in isopycnal mode (results were similar and are not presented here).
%(select results are shown in the supplementary material).

@@

\subsection{Breaking bonds}
The numerical bonds in the iceberg model enable the tabular iceberg to retain its shape. This is demonstrated by comparing the Control simulation to an identical simulation where all numerical bonds have been removed (Figure \ref{fig:Drift_berg}, movie S5). In the bond-free simulation, the ice elements disperse and the calved iceberg quickly loses its original structure. This bond-free simulation does not adequately represent tabular iceberg, which can move long distances through the ocean as a cohesive unit. This result motivates the inclusion of bonds in the iceberg model, even though they are more computationally expensive than traditional point-iceberg models. 
%The model behavior in the bond-free simulations is similar to an ice-shelf disintegration event.%, suggesting that 
%Breaking multiple bonds simultaneously could be used to simulate ice shelf disintegration events. \textcolor{red}{??? either add something more or remove this sentence ???}

@@

The  wave radiation force $(\vec{F}_{R})$ is given by
\begin{equation}
\vec{F}_{R} = \rho_{o} c_{r} g a  \textrm{ } \frac{WL}{W+L} \frac{\vec{v}_{a}}{|\vec{v}_{a} |} \textrm{ min}(a,F) 
\end{equation}
where g is the acceleration due to gravity, a is the wave amplitude empirically related to the wind speed by  $a=0.010125|\vec{v}_{a} -  \vec{v}_{o}| $, and $c_{wd}$ is the wave drag coefficient defined as 
\begin{equation}
c_{wd}=0.06 \textrm{ min} \Bigg( \textrm{ max} \Bigg[0,\frac{L-L_{c}}{L_{t}-L_{c} } \Bigg]  , 1    \Bigg), 
\end{equation}
where $L_{w}=0.32|\vec{v}_{a} -  \vec{v}_{o}|^{2}$ is an empirical wave length, $L_{c}=0.125 L_{w}$ is the cutoff length, and  $L_{t}=0.25 L_{w}$ is the upper limit.

@@

\subsection{Melt rate parametrization}
As discussed in Section 2.5, unbounded ice elements in the iceberg model decay according to parameterizations for iceberg decay typically used in iceberg drift models \citep{Martin2010}, while ice elements within larger ice structures have only a basal melt given by the three equation model \citep{Holland1999} .

For unbonded ice elements, the element thickness decays due to basal melt at a rate $M_{b}$, while the length and width of the elements decay as a result of melt due to wave erosion, $M_{e}$, and melt due to buoyant convection, $M_{v}$. Following \citet{Gladstone2001} and \citet{Martin2010}, the basal melt rate, wave erosion melt rate, and buoyant convection melt rate are parameterized by 
\begin{equation}
M_{b} = 0.58 | \vec{v} - \vec{v}_{0}|^{0.8} \frac{\tilde{T}_{0} - \tilde{T}}{L^{0.2}}
\end{equation}
\begin{equation}
M_{e} = \frac{1}{12} S_{s} \Bigg(  1 + \cos{[\pi A_{i}^{3}]}  \Bigg) \Bigg( \tilde{T}_{0} +2 \Bigg ),
\end{equation}
\begin{equation}
M_{v} =  \Bigg(7.62 \times 10^{-3}  \Bigg )   \tilde{T}_{0} + \Bigg( 1.29 \times 10^{-3} \Bigg)\tilde{T}_{0}^{2}.
\end{equation}
$\tilde{T}$ is the effective iceberg temperature and is set to $\tilde{T}=-4^{\circ}$C, $\tilde{T}_{0}$ is the temperature at the top of the ocean,  $A_{i}$ is the sea-ice area fraction, and $S_{s}$ is the sea state, which is given by the Beaufort scale
\begin{equation}
S_{s} =  \frac{2}{3}  | \vec{u}_{a} - \vec{u}_{o} |^{\frac{1}{2}}  +\frac{1}{10}   | \vec{u}_{a} -\vec{u}_{o} |
\end{equation}
All three melt rates are in units of meters per day.

@@

%The point-particle iceberg melt rates and the 3 equation model melt rates are both calculated using the ocean mixed-layer properties, which are passed from the ocean model to the iceberg model.

@@

\section{Appendix B}
\subsection{Modified Verlet Algorithm}
The model uses a version velocity Verlet time-stepping algorithm, which has been modified to allow part of the forcing to be calculated implicitly. The traditional velocity Verlet algorithm is commonly used in molecular dynamics, as it is simple to implement, second order accurate and computationally efficient \citep{Swope1982, Omelyan2002}.
Here we modify the traditional scheme to allow for the drag forces to be modeled implicitly, which prevents large accelerations for element's whose mass approaches zero. 
To do this, we include both an implicit and explicit acceleration, $a=a^{exp} +a^{imp}$. The explicit acceleration, $a^{exp}$, includes all forcing terms which depend only on the previous time step and the current position, while the implicit acceleration, $a^{imp}$, includes forcing terms which depend on the velocity at the current time step (in particular the drag and Coriolis forces).

@@

Using a time step of $\Delta t$, and subscripts to denote the time step (so that $t_{n+1}= t_{n} + \Delta t$), the modified velocity Verlet scheme can be written as:\\
1) Calculate updated position: $ x_{n+1} = x_{n} + u_{n} \Delta t  + \displaystyle{\frac{ \Delta t^{2}}{2}} \Bigg(a^{exp}_{n} + a^{imp}_{n} \Bigg) $.\\
2) Calculate $a^{exp}_{n+1}$\\
% (which is a function of $x_{n+1}$, and other quantities measured at time $t_{n}$). \\
3) Calculate $a^{imp}_{n+1}$ and $u_{n+1} = u_{n} + \displaystyle{\frac{\Delta t}{2}} \Bigg(a^{exp}_{n} + a^{exp}_{n+1} \Bigg) +  (\Delta t) a^{imp}_{n+1}$\\
%  (which is a function of $u_{n+1}$, $x_{n+1}$, and other quantities measured at time $t_{n}$)
This scheme reduces to the traditional velocity Verlet when $a^{imp}$ is set to zero. Note that $a^{exp}_{n+1} = a^{exp}_{n+1}(x_{n+1}, t_{n})$ is an explicit function of  $x_{n+1}$ and other quantities evaluated at time $t_{n}$, while $a^{imp}_{n+1}=a^{imp}_{n+1}(u_{n+1}, x_{n+1}, t_{n})$ additionally depends on $u_{n+1}$, and needs to be solved implicitly. For this reason in step three, $a^{imp}_{n+1}$ and $u_{n+1}$ need to be solved simultaneously, as described in the next subsection.

@@

\section{Appendix C}
\textbf{Connecting bonds across processor boundaries}\\
Since the model is parallelized across multiple processors, it often happens that two elements on different processes are bonded together. Keeping track of numerical bonds across processor boundaries requires a lot of book keeping. In this section we describe the how the model handles bonds across processor boundaries.

@@

\clearpage
\begin{table}[h!]
\begin{center}
\begin{tabular}{ |c c c c| }
\hline
Parameter & Symbol & Value &  Unit \\
\hline
Domain Length &  $L_{x}$ & 80  & km \\
Domain Width &  $L_{y}$ & 480  & km \\
Horizontal Resolution &  $\Delta x$ & 2  & km \\
Number of vertical layers &  $N_{l}$ & 72  & non-dim\\
Horizontal Viscosity &  $\nu_{H}$ & 6.0 & $m^{2}$ $s^{-1}$ \\
Diapycnal Viscosity &  $\nu_{V} $ & $10^{-3}$ & $m^{2}$ $s^{-1}$\\
Horizontal Diffusivity & $ \epsilon_{H}$ & 1.0 & $m^{2}$ $s^{-1}$ \\
Diapycnal Diffusivity &  $\epsilon_{V} $ & $5 \times 10^{-5}$ & $m^{2}$ $s^{-1}$ \\
Initial Surface Temperature &  $ T_{t}$& -1.9   & $ ^{o} C$\\
Initial Bottom Temperature &  $ T_{b}$&  1.0  & $ ^{o} C$\\
Initial Surface Salinity &  $ S_{t}$& 33.8   & psu\\
Initial Bottom Salinity &  $ S_{b}$&  34.7  & psu\\
Maximum Ocean depth &  $H_{ocean} $ & 720 & m \\
Relaxation Time of Sponge Layer &  $T_{sponge}$ & 0.1  & days\\
Length of Sponge Layer &  $L_{sponge}$ & 10  & km\\
%Background Frictional Velocity Below Ice &  $u^{*}_{bg}$ & 0.0  & 104 \\
%Drag Coefficient Below Ice&  c_{d} & $0.0025$  & (non dim)\\
%Class 9 &  $3.9\times 10^{11}$ & 1659 & 104 \\
%Time Step for Static Shelf Experiment & $dt_{Static}$ &  1000  & s\\
Ocean and iceberg model time step & $dt$ &  10  & s\\
Elastic interactive force spring constant & $ \kappa_{e} $ &  $10^{-5}$  & kg $\textrm{s}^{-2}$ \\
%Class 1 &  $8.8\times 10^{7}$ & 60  & 0.0026 & 40 & 2000 \\
%Class 2 &  $4.1\times 10^{8}$ & 100 & 0.0072 & 67 & 200\\
%Class 3 &  $3.3\times 10^{9}$ & 200  & 0.029 &133 & 50 \\
%Class 4 &  $1.8\times 10^{10}$& 350 &  0.12 &175  & 20 \\
%Class 5 &  $3.8\times 10^{10}$ & 500  & 0.18 &250 & 10\\
%Class 6 &  $7.5\times 10^{10}$ & 700   & 0.35 &250 & 5\\
%Class 7 &  $1.2\times 10^{11}$ & 900  & 0.56 &250  & 2 \\
%Class 8 &  $2.2\times 10^{11}$ & 1200  & 1.0&250 & 1 \\
%Class 9 &  $3.9\times 10^{11}$ & 1600 & 1.8&250 & 1  \\
%Class 10 & $7.4\times 10^{11}$ &  2200 & 3.5 &250  &1\\
\hline
\end{tabular}
\caption{Parameters used in the model. The ocean model parameters are as described in \citet{Asay-Davis2016}}.
% Note that the areas of the iceberg classes (calculated from the iceberg lengths and masses given in \citep{Gladstone2001}) are not monotonically increasing.
\label{table:parameters2}
\end{center}
\end{table}

@@

%\begin{figure}
%\begin{center}
%\includegraphics[width=0.99\textwidth]{/Users/alon/Desktop/files/Icebergs_clusters/Towards_Publication/Tech_paper/Github_stuff/Tech-paper/Figures/snapshots_ALE_z_Mixed_Melt_Collapse_spread_area.png}
%\caption{ {Snapshots of the fraction of ice cover in the iceberg model tabular iceberg calving simulation. Snapshots are taken (a) 7, (b) 15, and (c) 30 days after calving. The dashed line in panel (c) shows the location of the vertical transects shown in Figures \ref{fig:Velocity_section_Collapse} and \ref{fig:Temperature_section_Collapse}}. \label{fig:Spread_area}}
%\end{center}
%\end{figure}
 %\clearpage

\begin{figure}
\begin{center}
\includegraphics[width=0.99\textwidth]{/Users/alon/Desktop/files/Icebergs_clusters/Towards_Publication/Tech_paper/Github_stuff/Tech-paper/Figures/snapshots_ALE_z_Mixed_Melt_Collapse_temp.png}
\caption{ {Snapshots of the sea surface temperature in the tabular iceberg calving simulation. Snapshots are taken (a) 7, (b) 15, and (c) 30 days after calving. Grid cells with ice mass > $10^{4}$ kg are plotted in white, with grey shading indicating thinner ice.  The dashed line in panel (c) shows the location of the vertical transects shown in Figures \ref{fig:Temperature_section_Collapse} and \ref{fig:Velocity_section_Collapse}.} 
\label{fig:SST_Collapse}}
\end{center}
\end{figure}
 \clearpage
